---
output: html_document
---
# Data Manipulation and Plotting in R: the dplyr and ggplot2 packages

```{r, echo=FALSE, purl=FALSE, message = FALSE}
setwd("/Users/brooklebee/Desktop/iron-data-workshop/")
iron <- read.csv("data/iron.csv")
iron$station_id <- as.factor(iron$station_id)
```

------------

> ## Learning Objectives
>
> By the end of this lesson the learner will:
>
> * Know the six basic data manipulation 'verbs' in the dplyr package
> * Be able to select subsets of columns from a dataframe, and filter rows according to a condition(s)
> * Use the 'pipe' operator to link together a sequence of dplyr verbs
> * Be able to create new columns of data by applying functions to existing columns using the 'mutate' command
> * Know how to export a dataframe to a csv file using write.csv

------------

# Data Manipulation using dplyr

Bracket subsetting is handy, but it can be cumbersome and difficult to read,
especially for complicated operations. Enter `dplyr`. `dplyr` is a package for
making data manipulation easier. As you'll see, by data 'manipulation' we mean 
organizing and re-organizing your data so that you can better understand and 
present them, not any less ethical sense of the word.

Packages in R are basically sets of additional functions that let you do more
stuff. The functions we've been using so far, like `str()` or `data.frame()`,
come built into R; packages give you access to more of them. Before you use a
package for the first time you need to install it on your machine, and then you
should import it in every subsequent R session when you need it.

```{r, eval = FALSE, purl = FALSE}
install.packages("dplyr")
```

You might get asked to choose a CRAN mirror -- this is basically asking you to
choose a site to download the package from. The choice doesn't matter too much;
we recommend the RStudio mirror.

```{r, message = FALSE, purl = FALSE}
library("dplyr")    ## load the package
```

## What is `dplyr`?

The package `dplyr` provides easy tools for the most common data manipulation
tasks. It is built to work directly with data frames. The thinking behind it was
largely inspired by the package `plyr` which has been in use for some time but
suffered from being slow in some cases.` dplyr` addresses this by porting much
of the computation to C++. An additional feature is the ability to work directly
with data stored in an external database. The benefits of doing this are
that the data can be managed natively in a relational database, queries can be
conducted on that database, and only the results of the query returned.

This addresses a common problem with R in that all operations are conducted in
memory and thus the amount of data you can work with is limited by available
memory. The database connections essentially remove that limitation in that you
can have a database of many 100s GB, conduct queries on it directly, and pull
back just what you need for analysis in R.

To learn more about `dplyr` after the workshop, you may want to check out this
[handy dplyr cheatsheet](http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).


## Selecting columns and filtering rows

We're going to learn some of the most common `dplyr` functions: `select()`,
`filter()`, `mutate()`, `group_by()`, and `summarize()`. To select columns of a
data frame, use `select()`. The first argument to this function is the data
frame (`iron`), and the subsequent arguments are the columns to keep.

```{r, results = 'hide', purl = FALSE}
iron_select <- select(iron, station_id, datetime_utc, rain_inches, air_temp_f)
head(iron_select)
```

To choose rows, use `filter()`:

```{r, purl = FALSE}
iron_station4 <- filter(iron, station_id == 4)
head(iron_station4)
```

## Pipes

But what if you wanted to select and filter at the same time? There are three
ways to do this: use intermediate steps, nested functions, or pipes. With the
intermediate steps, you essentially create a temporary data frame and use that
as input to the next function. This can clutter up your workspace with lots of
objects. You can also nest functions (i.e. one function inside of another).
This is handy, but can be difficult to read if too many functions are nested as
the process from inside out. The last option, pipes, are a fairly recent
addition to R. Pipes let you take the output of one function and send it
directly to the next, which is useful when you need to do many things to the same
data set.  Pipes in R look like `%>%` and are made available via the `magrittr`
package installed as part of `dplyr`.

```{r, purl = FALSE, eval = FALSE}
iron %>%
  filter(rain_inches > 0) %>%
  select(station_id, datetime_utc, rain_inches)
```

In the above we use the pipe to send the `iron` data set first through
`filter`, to keep rows where `rain` was greater than 0, and then through `select`
to keep the `station_id`, `datetime_utc`, and `rain_inches` columns. When the data frame 
is being passed to the `filter()` and `select()` functions through a pipe, we don't need 
to include it as an argument to these functions anymore.

If we wanted to create a new object with this smaller version of the data we
could do so by assigning it a new name:

```{r, purl = FALSE}
iron_rain <- iron %>%
  filter(rain_inches > 0) %>%
  select(station_id, datetime_utc, rain_inches)

head(iron_rain)
```

Note that the final data frame is the leftmost part of this expression.

> ### Challenge {.challenge}
>
>  Using pipes, subset the data to include rows collected at stations 1 or 2,
>  and retain the columns `date`, `time_utc`, and `air_temp_f.`

<!---
```{r, eval=FALSE, purl=FALSE}
## Answer
iron %>%
    filter(station_id == 1 | station_id == 2) %>%
    select(date, time_utc, air_temp_f)
```
--->


### Mutate

Frequently you'll want to create new columns based on the values in existing
columns, for example to do unit conversions, or find the ratio of values in two
columns. For this we'll use `mutate()`.

To create a new column of rain in centimeters:

```{r, purl = FALSE, eval = FALSE}
iron %>%
  mutate(rain_cm = rain_inches / 2.54)
```

If this runs off your screen and you just want to see the first few rows, you
can use a pipe to view the `head()` of the data (pipes work with non-dplyr
functions too, as long as the `dplyr` or `magrittr` packages are loaded).

```{r, purl = FALSE}
iron %>%
  mutate(rain_cm = rain_inches / 2.54) %>%
  head
```

The first few rows all have 0 rain, so if we wanted to remove those we could
insert a `filter()` in this chain:

```{r, purl = FALSE}
iron %>%
  filter(rain_inches != 0) %>%
  mutate(rain_cm = rain_inches / 2.54) %>%
  head
```

The `!` symbol negates a statement, so `!=` means 'not equal'. Sometimes you might want to 
filter out NAs. `is.na()` is a function that determines whether something is or is not an 
`NA`. You could use `!is.na()` to ask for everything that is not an `NA`.

> ### Challenge {.challenge}
>
>  Create a new dataframe from the iron data that meets the following
>  criteria: contains only the `station_id` column and a column that contains
>  values that are ten times the `water_content_8inch` values (e.g. a new column
>  `water_content_8inch_ten`). In this `water_content_8inch_ten` column, all 
>  values are greater than 2.
>
>  **Hint**: think about how the commands should be ordered to produce this data frame!

<!---
```{r, eval=FALSE, purl=FALSE}
## Answer
iron_wc_8in_ten <- iron %>%
    mutate(water_content_8inch_ten = water_content_8inch * 10) %>%
    filter(water_content_8inch_ten > 2) %>%
    select(station_id, water_content_8inch_ten)
```
--->

   
### Split-apply-combine data analysis and the summarize() function

Many data analysis tasks can be approached using the "split-apply-combine"
paradigm: split the data into groups, apply some analysis to each group, and
then combine the results. `dplyr` makes this very easy through the use of the
`group_by()` function.


#### The `summarize()` function

`group_by()` is often used together with `summarize()` which collapses each
group into a single-row summary of that group.  `group_by()` takes as argument
the column names that contain the **categorical** variables for which you want
to calculate the summary statistics. So to view mean the `air_temp_f` by station:

```{r, purl = FALSE}
iron %>%
  group_by(station_id) %>%
  summarize(mean_temp = mean(air_temp_f, na.rm = TRUE))
```

You can group by multiple columns too, if you have another factor. Let's make a variable 
'rained' to represent whether non-zero rainfall was measured, then use dplyr to group by 
station_id and this new variable. For the first part, we're going to use an 
if...else statement, which we will not cover in detail but which may prove useful. I 
recommend checking out [this resource](https://www.programiz.com/r-programming/if-else-statement) for more information. 

```{r, purl = FALSE}
iron$rained <- ifelse(iron$rain_inches == 0, "N", "Y")
class(iron$rained)
iron$rained <- as.factor(iron$rained)

iron %>%
  group_by(station_id, rained) %>%
  summarize(mean_temp = mean(air_temp_f, na.rm = TRUE))
```

If you are working with a dataset with missing data, 'na.rm = TRUE' can become important. 
Some R functions (like min()) will output `NaN` (which refers to "Not a Number") if given 
a set of numeric values that contain `NA`. To avoid this, we could remove the missing 
values before we attempt to calculate the summary statistics with `!is.na()`. Because the 
missing values are removed, we can omit `na.rm=TRUE` when computing the mean:

```{r, purl = FALSE}
iron %>%
  filter(!is.na(air_temp_f)) %>%
  group_by(station_id, rained) %>%
  summarize(mean_temp = mean(air_temp_f))
```

You may also have noticed, that the output from these calls don't run off the
screen anymore. That's because `dplyr` has changed our `data.frame` to a
`tbl_df`. This is a data structure that's very similar to a data frame; for our
purposes the only difference is that it won't automatically show tons of data
going off the screen, while displaying the data type for each column under its
name. If you want to display more data on the screen, you can add the `print()`
function at the end with the argument `n` specifying the number of rows to
display:

```{r, purl = FALSE}
iron %>%
  filter(!is.na(air_temp_f)) %>%
  group_by(station_id, rained) %>%
  summarize(mean_temp = mean(air_temp_f)) %>%
  print(n=15)
```

Once the data is grouped, you can also summarize multiple variables at the same
time (and not necessarily on the same variable). For instance, we could add a
column indicating the minimum temperature for each station:

```{r, purl = FALSE}
iron %>%
  filter(!is.na(air_temp_f)) %>%
  group_by(station_id) %>%
  summarize(mean_temp = mean(air_temp_f),
            min_temp = min(air_temp_f))
```

#### Tallying

When working with data, it is also common to want to know the number of
observations found for each factor or combination of factors. For this, `dplyr`
provides `tally()`. For example, if we wanted to group by whether it rained and find the
number of rows of data for rained Y versus rained N, we would do:

```{r, purl = FALSE}
iron %>%
  group_by(rained) %>%
  tally()
```

Here, `tally()` is the action applied to the groups created by `group_by()` and
counts the total number of records for each category.

> ### Challenge {.challenge}
>
> How many measurements were made at each station for rained Y versus rained N?

<!---
```{r, echo=FALSE, purl=FALSE}
## Answer
iron %>%
    group_by(station_id, rained) %>%
    tally()
```
--->

> ### Challenge {.challenge}
>
> Use `group_by()` and `summarize()` to find the mean, min, and max air
> temperature for each station (using `station_id`).

<!---
```{r, echo=FALSE, purl=FALSE}
## Answer
iron %>%
    filter(!is.na(air_temp_f)) %>%
    group_by(station_id) %>%
    summarize(
        mean_temp = mean(air_temp_f),
        min_temp = min(air_temp_f),
        max_temp = max(air_temp_f)
    )
```
--->

> ### Challenge {.challenge}
>
> What was the most rain measured on any one day at any station?

<!---
## Answer
```{r, echo=FALSE, purl=FALSE}
maxrain <- iron %>%
    filter(!is.na(rain_inches)) %>%
    group_by(station_id, date) %>%
    summarize(daily_rain_in = sum(rain_inches)) %>%
    filter(daily_rain_in == max(daily_rain_in)) %>%
    arrange(desc(daily_rain_in))
```
--->


# Exporting data

Now that you have learned how to use `dplyr` to extract the information you need
from the raw data, or to summarize your raw data, you may want to export these
new datasets to share them with your collaborators or for archival.

Similarly to the `read.csv()` function used to read in CSV into R, there is a
`write.csv()` function that generates CSV files from data frames.

Before using it, we are going to create a new folder, `data_output` in our
working directory that will store this generated dataset. We don't want to write
generated datasets in the same directory as our raw data. It's good practice to
keep them separate. The `data` folder should only contain the raw, unaltered
data, and should be left alone to make sure we don't delete or modify it; on the
other end the content of `data_output` directory will be generated by our
script, and we know that we can delete the files it contains because we have the
script that can re-generate these files.

We can save our dataframe as a CSV file in our `data_output`
folder. By default, `write.csv()` includes a column with row names (in our case
the names are just the row numbers), so we need to add `row.names = FALSE` so
they are not included:

```{r, purl=FALSE, eval=FALSE}
write.csv(iron, file="data_output/iron_unchanged.csv",
          row.names=FALSE)
```


### Quick break to talk a bit more about dates and datetimes in R

As we've already discussed, dates and times can be difficult to represent correctly when
communicating with computers (and other people, too!). We've recommended some ways to 
store dates that can help with this, including some standard formats.

Here we're going to briefly introduce another way to deal with dates AND times in R 
using POSIX format. For more information and other approaches, check out 
[this helpful guide](https://www.stat.berkeley.edu/~s133/dates.html), from which this 
section borrows heavily.

POSIX stands for "portable operating system interface" and is used by many operating 
systems, including UNIX systems. Dates stored in the POSIX format are date/time values 
and allow modification of time zones. POSIX date classes store times to the nearest 
second, which can be useful if you have data at that scale.

There are two POSIX date/time classes, which differ in the way that the values are stored 
internally. The POSIXct class stores date/time values as the number of seconds since 
January 1, 1970, while the POSIXlt class stores them as a list with elements for 
second, minute, hour, day, month, and year, among others. Unless you need the list nature 
of the POSIXlt class, the POSIXct class is the usual choice for storing dates in R. 
The ggplot2 plotting package that we will introduce next uses the POSIXct class.

We've generated a column 'datetime_utc' in the iron.csv dataset that is already in the
default input format for POSIX dates: the year, followed by the month and day, separated 
by slashes or dashes. For date/time values, the date may be followed by white space (e.g.
space or tab) and a time in the form hour:minutes:seconds or hour:minutes, which then 
may be followed by white space and the time zone. Here are some examples of valid POSIX 
inputs:

1915/6/16
2005-06-24 11:25
1990/2/17 12:20:05
2012-7-31 12:20:05 MST

What class is the iron$datetime_utc column in after you import it into R? If it's not already in POSIX format, you'll need to modify it using the as.POSIX() function.

```{r, purl=FALSE}
## Formate datetime_utc appropriately.
class(iron$datetime_utc)
iron$datetime_utc <- as.POSIXct(iron$datetime_utc, tz="UTC")
class(iron$datetime_utc)

```

Great! That sets us up to start plotting values over time.


# Creating beautiful plots with ggplot2

#### Disclaimer

We will be using the functions in the ggplot2 package. R has powerful built-in
plotting capabilities, but for this exercise, we will be using the ggplot2
package, which facilitates the creation of highly-informative plots of
structured data.

> ### Learning Objectives
>
> By the end of this lesson the learner will:
>
> * Be able to create a ggplot object
> * Be able to set universal plot settings
> * Be able to modify an existing ggplot object
> * Be able to change the aesthetics of a plot such as colour
> * Be able to edit the axis labels 
> * Know how to use a step-by-step approach to build complex plots
> * Be able to create, scatter plots, box plots and time series plots
> * Use the facet_ commands to create a collection of plots splitting the data by a factor variable
> * Be able to create customized plot styles to meet their needs

We start by loading the required packages.

```{r, purl=FALSE}
# plotting package
library(ggplot2)

# modern data frame manipulations
library(dplyr)
```

## Plotting with ggplot2

We will all make the same plot using the `ggplot2` package.

`ggplot2` is a plotting package that can create complex plots iteratively
from data in a dataframe. It uses default settings, which help creating
publication quality plots with a minimal amount of settings and tweaking.

ggplot graphics are built step by step by adding new elements.

To build a ggplot we need to:

- bind the plot to a specific data frame using the `data` argument

```{r, eval=FALSE}
ggplot(data = iron)
```

- define aesthetics (`aes`), by selecting the variables to be plotted and the variables to define the presentation
     such as plotting size, shape color, etc.,

```{r, eval=FALSE}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches))
```

- add `geoms` -- graphical representation of the data in the plot (points,
     lines, bars). To add a geom to the plot use `+` operator:

```{r first-ggplot, eval=FALSE}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches)) + geom_point()
```

The `+` in the `ggplot2` package is particularly useful because it allows you
to modify existing `ggplot` objects. This means you can easily set up plot
"templates" and conveniently explore different types of plots, so the above
plot can also be generated with code like this:

```{r, first-ggplot-with-plus, eval=FALSE}
# Create
iron_plot <- ggplot(data = iron, aes(x = air_temp_f, y = rain_inches))

# Draw the plot
iron_plot + geom_point()
```

Notes:

- Anything you put in the `ggplot()` function can be seen by any geom layers
  that you add (i.e., these are universal plot settings). This includes the x and
  y axis you set up in `aes()`.
- You can also specify aesthetics for a given geom independently of the
  aesthetics defined globally in the `ggplot()` function.


## Building your plots iteratively

Building plots with ggplot is typically an iterative process. We start by
defining the dataset we'll use, lay the axes, and choose a geom.

```{r create-ggplot-object}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches)) +
    geom_point()
```

Then, we start modifying this plot to extract more information from it. For
instance, we can add transparency (alpha) to avoid overplotting.

```{r adding-transparency}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches)) +
    geom_point(alpha = 0.5)
```

We can also add colors for all the points.

```{r adding-colors}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches)) +
    geom_point(alpha = 0.5, color = "blue")
```

Or color each station in the plot differently.

```{r color-by-station}
ggplot(data = iron, aes(x = air_temp_f, y = rain_inches)) +
    geom_point(alpha = 0.5, aes(color=station_id))
```

## Boxplot

Visualising the distribution of air temperatures at each station.

```{r boxplot}
ggplot(data = iron, aes(x = station_id, y = air_temp_f)) +
    geom_boxplot()
```

By adding points to boxplot, we can have a better idea of the number of
measurements and of their distribution:

```{r boxplot-with-points}
ggplot(data = iron, aes(x = station_id, y = air_temp_f)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```

Notice how the boxplot layer is behind the jitter layer? What could you change in the code to put the boxplot in front of the points?

> ### Challenges
>
> Boxplots are useful summaries, but hide the *shape* of the distribution. For
> example, if there is a bimodal distribution, this would not be observed with a
> boxplot. An alternative to the boxplot is the violin plot (sometimes known as a
> beanplot), where the shape (of the density of points) is drawn.
>
> - Replace the box plot with a violin plot; see `geom_violin()`
>
>
> In many types of data, it is important to consider the *scale* of the
> observations.  For example, it may be worth changing the scale of the axis to
> better distribute the observations in the space of the plot.  Changing the scale
> of the axes is done similarly to adding/modifying other components (i.e., by
> incrementally adding commands).
>
> - Represent rainfall on the log10 scale; see `scale_y_log10()`.
>
> - Create boxplot for `water_content_2inch`.
>
> - Add color to the datapoints on your boxplot according to the time at which the measurement was taken (`time_utc`).
>
> Hint: Check the class for `time_utc`. Consider changing the class of `time_utc` from factor to numeric. Why does this change how R makes the graph? 


## Plotting data across time

Let's start by looking at how air temperature changes over time.

```{r first-time-series}
ggplot(data = iron, aes(x = datetime_utc, y = air_temp_f)) +
     geom_line()
```

Huh, looks like there's a lot of noise there, both due to daily variation and also maybe differences between stations. Let's calculate the mean temperature for each day at each station. To do that we need to group data first and count records within each group, with our old friend dplyr.

```{r}
daily_air_temp <- iron %>%
                group_by(date, station_id) %>%
    			summarize(mean_temp = mean(air_temp_f))
```

Data across time can be visualised as a line plot with time on the x-axis and counts
on the y-axis.

```{r mean-time-series}
ggplot(data = daily_air_temp, aes(x = date, y = mean_temp)) +
     geom_line()
```

Unfortunately this looks really weird, for two reasons:

1. We need to treat the dates as dates, rather than integers.

 ```{r fix-date-class}
class(daily_air_temp$date)
daily_air_temp$date <- as.Date(as.character(daily_air_temp$date), format='%Y%m%d')
```

2. We are plotting data for all the stations together. We need to tell ggplot to draw a 
line for each station by modifying the aesthetic function to include `group = station_id`.

```{r time-series-by-station}
ggplot(data = daily_air_temp, aes(x = date, y = mean_temp, group = station_id)) +
    geom_line()
```

We will be able to distinguish stations in the plot if we add colors.

```{r time-series-with-colors}
ggplot(data = daily_air_temp, aes(x = date, y = mean_temp, group = station_id, colour = station_id)) +
    geom_line()
```

## Faceting

ggplot has a special technique called *faceting* that allows to split one plot
into multiple plots based on a factor included in the dataset. We will use it to
make one plot for a time series for each station.

```{r first-facet}
ggplot(data = daily_air_temp, aes(x = date, y = mean_temp, group = station_id, colour = station_id)) +
    geom_line() +
    facet_wrap(~ station_id)
```

Now we would like to split the line in each plot by whether rain was detected at the same 
time. To do that we need to calculate means in a data frame grouped by date, station_id, 
and the variable `rained` we created above.

```{r}
air_temp_by_rain <- iron %>%
                group_by(date, station_id, rained) %>%
    			summarize(mean_temp = mean(air_temp_f))
    			
air_temp_by_rain$date <- as.Date(as.character(air_temp_by_rain$date), format='%Y%m%d')

```

We can now make the faceted plot splitting further by whether it rained (at each station):

```{r facet-by-station-and-rain}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = station_id)) +
     geom_line() +
     facet_wrap(~ station_id)
```

Usually plots with white background look more readable when printed.  We can set
the background to white using the function `theme_bw()`. Additionally you can also remove 
the grid. 

```{r facet-by-station-and-rain-white-bg}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = station_id)) +
     geom_line() +
     facet_wrap(~ station_id) +
     theme_bw() +
     theme(panel.grid.major.x = element_blank(), 
	   panel.grid.minor.x = element_blank(),
	   panel.grid.major.y = element_blank(),
	   panel.grid.minor.y = element_blank())
```

To make the plot easier to read, we can color by whether it rained instead of station ID 
(stations are already in separate plots, so we don't need to distinguish them further).

```{r facet-by-station-and-rain-colored}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = rained)) +
    geom_line() +
    facet_wrap(~ station_id) +
    theme_bw()
```


## Challenge

> Use what you just learned to create a plot that depicts how the minimum 8-inch water 
> content of each station changes over time.

<!-- Answer
```{r average-8in-water-timeseries}
min_8in_water <- iron %>%
                group_by(date, station_id) %>%
    			summarize(min_content_8in = min(water_content_8inch, na.rm=TRUE))
    			
# Have to deal with the pesky dates.
min_8in_water$date <- as.Date(as.character(min_8in_water$date), format='%Y%m%d')

ggplot(data = min_8in_water, aes(x=date, y=min_content_8in, color = station_id, group = station_id)) +
    geom_line() +
    theme_bw()
```
-->


The `facet_wrap` geometry extracts plots into an arbitrary number of dimensions
to allow them to cleanly fit on one page. On the other hand, the `facet_grid`
geometry allows you to explicitly specify how you want your plots to be
arranged via formula notation (`rows ~ columns`; a `.` can be used as
a placeholder that indicates only one row or column).

Let's modify the previous plot to compare how the weights of male and females
has changed through time.

```{r mean-rain-time-facet-rain-rows}
## One column, facet by rows
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = station_id, color = station_id)) +
    geom_line() +
    facet_grid(rained ~ .)

```

```{r mean-rain-time-facet-rain-columns}
# One row, facet by column
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = station_id, color = station_id)) +
    geom_line() +
    facet_grid(. ~ rained)
```

## Customization

Take a look at the [ggplot2 cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/08/ggplot2-cheatsheet.pdf), and
think of ways to improve the plot. You can write down some of your ideas as
comments in the Etherpad.

Now, let's change names of axes to something more informative than 'date'
and 'mean_temp' and add a title to this figure:

```{r facet-by-station-and-rain-right-labels}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = rained)) +
    geom_line() +
    facet_wrap(~ station_id) +
    labs(title = 'Temperature At Four Stations',
         x = 'Date in 2016',
         y = 'Mean Daily Temperature (F)') +
    theme_bw()
```

The axes have more informative names, but their readability can be improved by
increasing the font size. While we are at it, we'll also change the font family:

```{r facet-by-station-and-rain-right-labels-xfont-size}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = rained)) +
    geom_line() +
    facet_wrap(~ station_id) +
    labs(title = 'Temperature At Four Stations',
         x = 'Date in 2016',
         y = 'Mean Daily Temperature (F)') +
    theme_bw() +
    theme(text=element_text(size=16, family="Arial"))
```

After our changes we notice that the values on the x-axis are not
properly readable. Let's change the orientation of the labels and adjust them
vertically and horizontally so they don't overlap. You can use a 90 degree
angle, or experiment to find the appropriate angle for diagonally oriented
labels.

```{r facet-by-station-and-rain-theme}
ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = rained)) +
    geom_line() +
    facet_wrap(~ station_id) +
    labs(title = 'Temperature At Four Stations',
         x = 'Date in 2016',
         y = 'Mean Daily Temperature (F)') +
    theme_bw() +
    theme(axis.text.x = element_text(colour="grey20", size=12, angle=90, hjust=.5, vjust=.5),
                        axis.text.y = element_text(colour="grey20", size=12),
          text=element_text(size=16, family="Arial"))
```

If you like the changes you created to the default theme, you can save them as
an object to easily apply them to other plots you may create:


```{r saving-themes}
arial_grey_theme <- theme(axis.text.x = element_text(colour="grey20", size=12, angle=90, hjust=.5, vjust=.5),
                          axis.text.y = element_text(colour="grey20", size=12),
                          text=element_text(size=16, family="Arial"))
ggplot(data = iron, aes(x = station_id, y = air_temp_f)) +
    geom_boxplot(alpha = 0) +
    arial_grey_theme
```

With all of this information in hand, please take another five minutes to either
improve one of the plots generated in this exercise or create a beautiful graph
of your own. Use the RStudio ggplot2 cheat sheet, which we linked earlier for
inspiration.

Here are some ideas:

* See if you can change thickness of the lines.
* Can you find a way to change the name of the legend? What about its labels?
* Use a different color palette (see http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)

After creating your plot, you can save it to a file in your favourite format (using device = either a device function (e.g. png), or one of "eps", "ps", "tex" (pictex), "pdf", "jpeg", "tiff", "png", "bmp", "svg" or "wmf" (windows only).
You can easily change the dimension (and its resolution) of your plot by
adjusting the appropriate arguments (`width`, `height` and `dpi`):

```{r ggsave-example, eval=FALSE}
myplot <- ggplot(data = air_temp_by_rain, aes(x = date, y = mean_temp, group = rained, color = rained)) +
    geom_line() +
    facet_wrap(~ station_id) +
    labs(title = 'Temperature At Four Stations',
         x = 'Date in 2016',
         y = 'Mean Daily Temperature (F)') +
    theme_bw() +
    theme(axis.text.x = element_text(colour="grey20", size=12, angle=90, hjust=.5, vjust=.5),
                        axis.text.y = element_text(colour="grey20", size=12),
          text=element_text(size=16, family="Arial"))
ggsave("name_of_file.png", myplot, width=15, height=10)
```


## Useful links for colors

[Using colorbrewer to choose colors, on the web](http://colorbrewer2.org/#type=qualitative&scheme=Paired&n=4)

[Colorbrewer within ggplot2](http://docs.ggplot2.org/current/scale_brewer.html)

[The RColorBrewer package](https://cran.r-project.org/web/packages/RColorBrewer/index.html)

[Colorblind-friendly figures](http://bconnelly.net/2013/10/creating-colorblind-friendly-figures/)
